Conversas Hiperligadas: Novo Paradigma de Comunicação e Colaboração, potenciado pela Tecnologia WebRTC

### 2016-05-09 tese2/10_related_work_hypermedia.tex
[ADD] template2
[RFT] rewriten model

\subsection{Hypermedia: more than words, more than images}
\label{hypermedia}

@@

  \ac{SMIL} is an \ac{XML} based language that defines twelve modules: \emph{Animation}, \emph{Content Control}, \emph{Layout}, \emph{Linking}, \emph{Media Objects}, \emph{SmilText}, \emph{Metainformation}, \emph{Structure}, \emph{Timing}, \emph{Time Manipulations}, \emph{State} and \emph{Transitions}.

@@

  \item The \textbf{Metainformation} module contains elements and attributes that allows describing the \ac{SMIL} document. For example, this module could be used to define movie details such as category, director, writers and cast.

  \item The \textbf{Structure} module defines the basic elements and attributes for structuring \ac{SMIL} content. This module defines a \emph{head} element that contains non temporal behavior information defined by  \emph{Metainformation}, \emph{Layout} and \emph{Content Control} modules. This module also defines the \emph{body} element, where all temporal related module information is contained.

@@

%RP explicar melhor. Não vejo como low-latency pode ser mau
  %HR troquei-me todo
  With the emergence of \ac{HTML}5, tags like \emph{video}, \emph{audio} and \emph{track} allow us to play video with multiple codecs, audio and subtitles in \ac{WebVTT} format. Another important tag is \emph{canvas} that allows drawing graphics with \emph{JavaScript} on a rectangle within a web page.

  For example, with \ac{API}s like WebGL\footnote{\url{http://khronos.org/webgl/}(accessed June 2, 2015).}, it is now possible to manipulate a three dimensional environment in the context of a hyper-call. Another example would be a collaborative spreadsheet using \ac{WebRTC}. With this, hyper-calls are not limited to only audio, image, text and video, but also interaction with complex graphical user interfaces that changes over time.
 %XXX
  %RP porque escolheste falar aogra de webgl se não falas antes?
%HR já troquei, o WebGL usa o canvas do HTML5, talvez seja este o local mais adequado

@@

  Back in 1995, \emph{flash}\footnote{\url{http://www.adobe.com/products/flash.html}(accessed June 2, 2015).} was developed for web-based animations. Introducing video support in 2002, \emph{flash} started to grow after that. Competitor's players, at that time were focused on playing video and audio, while \emph{flash} had vector graphics and focused on streaming \emph{on-demand} video across multiple platforms. \emph{VP6} was their choice of video codec, providing half the video size (in Byte) for the same quality and providing video quality adjusted to \emph{Internet} connection latency. 
  In 2010, \emph{Adobe Flash} was the most widely used applications for reproducing live broadcast and recorded video \cite{flashvideo}, it supports progressive video download using \ac{HTTP} and streaming using \ac{RTMP}.
  %RP ``was'' quando? Em 2010? Ontem?
  %HR exacto.

@@

  \ac{FMS} can stream video and audio to one or more subscribers by sending a separate copy for each subscriber. With \ac{RTMFP} it is possible to stream video directly between \emph{flash} players, allowing a publisher to break up a stream into pieces that can be cooperatively distributed in a P2P mesh. \ac{RTMFP} uses \ac{UDP} to speed packet delivery, which although it is not reliable, is well suited for video streaming. Like \ac{WebRTC}, \emph{flash} players also need to apply techniques like \ac{STUN} and \ac{TURN} for \ac{NAT} traversal.
  %RP as vezes fazes \emph{flash} e outras vezes não. Convém ser coerente.

>>>>

\section{Hypermedia: more than words, more than images}
\label{hypermedia}

@@

  \ac{SMIL} is an \ac{XML} based language that defines twelve modules: \emph{Animation}, \emph{Content Control}, \emph{Layout}, \emph{Linking}, \emph{Media Objects}, \emph{SmilText}, \emph{Meta Information}, \emph{Structure}, \emph{Timing}, \emph{Time Manipulations}, \emph{State} and \emph{Transitions}.

@@

  \item The \textbf{Meta Information} module contains elements and attributes that allows describing the \ac{SMIL} document. For example, this module could be used to define movie details such as category, director, writers and cast.

  \item The \textbf{Structure} module defines the basic elements and attributes for structuring \ac{SMIL} content. This module defines a \emph{head} element that contains non temporal behavior information defined by  \emph{Meta Information}, \emph{Layout} and \emph{Content Control} modules. This module also defines the \emph{body} element, where all temporal related module information is contained.

@@

  With the emergence of \ac{HTML}5, tags like \emph{video}, \emph{audio} and \emph{track} allow us to play video with multiple \emph{codecs}, audio and subtitles in \ac{WebVTT} format. Another important tag is \emph{canvas} that allows drawing graphics with \emph{JavaScript} on a rectangle within a web page.

  For example, with \ac{API}s like WebGL\footnote{\url{http://khronos.org/webgl/}(accessed June 2, 2015).}, it is now possible to manipulate a three dimensional environment in the context of a hyper-call. Another example would be a collaborative spreadsheet using \ac{WebRTC}. With this, hyper-calls are not limited to only audio, image, text and video, but also interaction with complex graphical user interfaces that changes over time.

@@

  Back in 1995, \emph{flash}\footnote{\url{http://www.adobe.com/products/flash.html}(accessed June 2, 2015).} was developed for web-based animations. Introducing video support in 2002, \emph{flash} started to grow after that. Competitor's players, at that time were focused on playing video and audio, while \emph{flash} had vector graphics and focused on streaming \emph{on-demand} video across multiple platforms. \emph{VP6} was their choice of video \emph{codec}, providing half the video size (in Byte) for the same quality and providing video quality adjusted to \emph{Internet} connection latency. 
  In 2010, \emph{Adobe Flash} was the most widely used applications for reproducing live broadcast and recorded video \cite{flashvideo}, it supports progressive video download using \ac{HTTP} and streaming using \ac{RTMP}.

@@

  \ac{FMS} can stream video and audio to one or more subscribers by sending a separate copy for each subscriber. With \ac{RTMFP} it is possible to stream video directly between \emph{flash} players, allowing a publisher to break up a stream into pieces that can be cooperatively distributed in a P2P mesh. \ac{RTMFP} uses \ac{UDP} to speed packet delivery, which although it is not reliable, is well suited for video streaming. Like \ac{WebRTC}, \emph{flash} players also need to apply techniques like \ac{STUN} and \ac{TURN} for \ac{NAT} traversal.

### 2016-05-09 tese2/13_implementation_model.tex
[RFT] rewriten model

%RP antes desta secção fala um introdução ao capítulo a explicar o que vai ser apresentado.

\section{Data Model}
The data model is a critical component of our solution, as a badly designed model can imply serious difficulties when implementing new features that are not part of the plans. During the course of this project, we had to redesign the model more than once in order to support new features.
%RP mudei ``most important'' para critical

%RP devias começar com um overview dos dados que é necessário guardar para a aplicação
%RP ao apresentar o modelo falas de coisas como mensagens para grupos, mas ainda não sabemos nada disso. Na arquitectura, devias falar da concepção, do modelo de comunicação. Que tens utilizadores, grupos, mensagens, chats, convites, etc...

@@

    Each \emph{MongoDB} document is represented by a \emph{JSON} object and, as a result, each document may have different attributes within a collection. Needless to say that, by following this approach we don't need to create the collections with a well predefined schema. In fact we don't need to define it at all. Although, for reasons of coherence and organization, we represent our database collections as they would have a predefined schema by following the same document structure.
    %RP a última frase é confusa. O que queres dizer? Reescreve

@@

    In respect to attributes nullability, \emph{MongoDB} does not enforce a document's attribute to have a non nullable value, although, for sake of good functionality, we perform those constraints validation programmatically and as such we also represent them in our schema.
    %RP nullable ou null?

@@

The identification number by itself is not sufficient to identify an object, as objects from different types can have the same identification number. In order to solve this problem, when an object is created, its correspondent data must contain the owner's object type. 
%RP mas o _id é apresentado com sendo uma chave primária na figura 4.2, logo única!

@@

@@

Entities that have a membership with a group can create more memberships by sharing an invite token or by specifying new group members.
%RP isto quer dizer que quem tem o token se pode juntar ao grupo? Explicar. O que impede alguém de o partihar com outros?

@@

In order to improve performance of queries over \emph{Relation} collection we have created indexes on the \emph{target} attribute and also on the pair composed by \emph{target} and \emph{\_id} attributes for finding and sorting the messages received by an entity more efficiently.
%RP pela justificação não consigo perceber para que serve a chave no par target e _id

@@

%RP explica melhor a diferença entre HyperContent e Time Annotation do ponto de vista do utilizador / funcionalidade proporcionada.

@@

During a conversation, users may allow sharing their web cameras. By doing so, their video is stored in recording chunks. Each chunk, described in Table \ref{table:recording}, represents an interval of time $T=\big[c^{start},c^{end}\big[$ of the video stream. It also contains the media's \ac{URL}, a reference to a group, an owner identification number and the correspondent \emph{WebSocket} session id.
    %RP ainda não tenho forma de perceber o que é o websocket session id. Nunca falaste de websockets antes, quanto mais disto! Nem do media URL, é gravado ou webrtc (dispositivo do utilizador)?

@@

In order to improve the performance of searching recording chunks and calculate which chunks are going to be reproduced, we have created indexes over the following sequences of attributes: 
\begin{itemize}
    \item{\emph{\{groupId,start,end\}} $\rightarrow$ for finding all available chunks for a given time} 
    \item{\emph{\{groupId,sessionId,end\}} $\rightarrow$ for finding intersecting chunks with a given time within a session}
    \item{\emph{\{groupId,owner,end\}} $\rightarrow$ for finding intersecting chunks with a given time within a user (for a different session) or group}
    \item{\emph{\{groupId,sessionId,\_id\}} $\rightarrow$ for finding chunks that follow a given \emph{id} within a session}
    \item{\emph{\{groupId,owner,\_id\}} $\rightarrow$ for finding chunks that follow a given \emph{id} within a user or group}
\end{itemize}
%RP os _id não são únicos? Para que servem os index que os incluem?
%RP para a intercepção não precisaste de usar o start, apenas o end?

Not less important, we have also created an index over the \emph{RecordingInterval}'s \emph{groupId} attribute for listing all intervals within a conference room. 
%RP olha outra coisa nova! Ainda não falaste de conference rooms!

>>>>

%RP antes desta secção fala um introdução ao capítulo a explicar o que vai ser apresentado.
%HR done

\section{Data Model}
The data model is a critical component of our solution, as a badly designed model can imply serious difficulties when implementing new features that are not part of the plans. During the course of this project, we had to redesign the model more than once in order to support new features.

%RP devias começar com um overview dos dados que é necessário guardar para a aplicação
%RP ao apresentar o modelo falas de coisas como mensagens para grupos, mas ainda não sabemos nada disso. Na arquitectura, devias falar da concepção, do modelo de comunicação. Que tens utilizadores, grupos, mensagens, chats, convites, etc...
%HR não meti na arquitectura mas adicionei a linha seguinte

In order to offer all the functionalities that we promise, some information about objects must be persistent such as users, groups,relations among users, group memberships, messages, hyper-content, recordings and collaborative editor state.

@@

    Each \emph{MongoDB} document is represented by a \emph{JSON} object and, as a result, each document may have different attributes within a collection. Needless to say that, by following this approach we don't need to create the collections with a well predefined schema. In fact we don't need to define it at all but, for reasons of coherence and organization, we represent our database collections as they would have a predefined schema by following the same document structure as we will present in this section.
    %RP a última frase é confusa. O que queres dizer? Reescreve
    %HR está melhor?

@@

    In respect to attributes nullability, \emph{MongoDB} does not enforce a document's attribute to have a not null value, although, for sake of good functionality, we perform those constraints validation programmatically and as such we also represent them in our schema.

@@

The owner's identification number by itself is not sufficient to identify an object, as objects from different types can have the same identification number. In order to solve this problem, when an object is created, its correspondent data must contain the owner's object type. 
%RP mas o _id é apresentado com sendo uma chave primária na figura 4.2, logo única!
%HR "The owner's identification number" corriji, não me referia ao "_id"

@@

A conference room, which is the environment where users communicate among their selves, is represented persistently by a group of participating entities (in general users but our system allow other types of entities).

@@

Entities that have a membership with a group can create more memberships by sharing an invite token or by specifying new group members. This invite token is used to create an invite \ac{URL} that if shared with other users allow them to join the group. Invite tokens can be deleted or regenerated with a different value.
%RP isto quer dizer que quem tem o token se pode juntar ao grupo? Explicar. O que impede alguém de o partihar com outros?
%HR done

@@

In order to query for recent messages for a given target, we use the oldest message's \emph{\_id}, which is sequential, in order to find messages with a newer \emph{\_id}.

In order to improve performance of queries over \emph{Relation} collection we have created indexes on the \emph{target} attribute and also on the pair composed by \emph{target} and \emph{\_id} attributes for finding and sorting the messages received by an entity more efficiently.

%RP pela justificação não consigo perceber para que serve a chave no par target e _id
%HR in order to find recent messages (não tinha explicado)

@@

In the point o view of a user, a time annotation is just a simple way to associate a time stamp to a topic in order to make it easier to find. On the other hand, the hyper-content is more complete than time annotations, except they are not visible on the time-line. In addition to time annotations, hyper-contents have a duration and content that can be superimposed to video.

%RP explica melhor a diferença entre HyperContent e Time Annotation do ponto de vista do utilizador / funcionalidade proporcionada.
%HR done

@@

During a conversation, users may allow sharing their web cameras. By doing so, their video is stored in recording chunks. Each chunk, described in Table \ref{table:recording}, represents an interval of time $T=\big[c^{start},c^{end}\big[$ of the video stream. It also contains the media's \ac{URL}, a reference to a group, an owner identification number and the correspondent \emph{WebSocket} session id.

    The media's \ac{URL} is the location where \emph{Kurento Repository} stores one chunk of video (including audio) which is used to playback on users demand.

    %RP ainda não tenho forma de perceber o que é o websocket session id. Nunca falaste de websockets antes, quanto mais disto! Nem do media URL, é gravado ou webrtc (dispositivo do utilizador)?

@@

In respect to find which chunks to play given a specified time (which is the current time in a user's perspective), we must select the chunk with an ending instant immediately after the current time, if the chunk's beginning is placed before the current time we have to play it instantaneously otherwise the chunk's beginning will occur after the current time and therefore we have to schedule the playback.

On the other hand, we use the fact that \emph{\_id}'s are sequential in order to determine adjacent chunks, for example, if we need to find the next chunk to play, we have to sort all elements by \emph{\_id} (which on \emph{MlongoDB} is already sorted by default) and select a chunk with an \emph{\_id} immediately after the current one.

In order to improve the performance of searching recording chunks and calculate which chunks are going to be reproduced, we have created indexes over the following sequences of attributes: 
\begin{itemize}
    \item{\emph{\{groupId,start,end\}} $\rightarrow$ for finding all available chunks for a given time} 
    \item{\emph{\{groupId,sessionId,end\}} $\rightarrow$ for finding chunks that ends after a given time within a session}
    \item{\emph{\{groupId,owner,end\}} $\rightarrow$ for finding chunks that ends after a given time within a user (for a different session) or group}
    \item{\emph{\{groupId,sessionId,\_id\}} $\rightarrow$ for finding chunks that follow a given \emph{id} within a session}
    \item{\emph{\{groupId,owner,\_id\}} $\rightarrow$ for finding chunks that follow a given \emph{id} within a user or group}
\end{itemize}
%RP os _id não são únicos? Para que servem os index que os incluem?
%HR done, for finding adjacent chunks

%RP para a intercepção não precisaste de usar o start, apenas o end?
%HR exactly, if there is no intersection taking into account the "end" at least I can schedulle the next player (after the end)

Not less important, we have also created an index over the \emph{RecordingInterval}'s \emph{groupId} attribute for listing all intervals within a conference room. 
%RP olha outra coisa nova! Ainda não falaste de conference rooms!
%HR "conference room" dá-se no contexto de um "group", era para não estar sempre a repetir. Qualquer das formas adicionei esse pormenor na seccao do grupo.

### 2016-05-09 tese2/implementation.tex
[ADD] change architecture
[ADD] deployment and test description
[ADD] chapter summary
[RFT] rewriten model

\input{13_implementation_model.tex}
\input{14_implementation_signaling.tex}
\input{15_implementation_recording.tex}

>>>>

In this chapter we are going to present our implementation choices, the difficulties that we have faced and the strategies available to overcome them.

To this end, we are going to present and explain our database model, how we have improved it and how this model can help us reach our goals.

Subsequently, we are going to describe our signaling protocol which will make it possible to use the \ac{WebRTC}'s functionalities in order to implement our system's streaming features.

With the model and signaling protocol defined we are prepared to show the multiple approaches we have followed in order to implement stream recording, their drawbacks and our choices.

Then are going to describe how we show hyper-content to users, namely the multiple ways to create hyper-content and the algorithm behind the user interface synchronization, the security flaws of our content displaying mechanism and a possible solution to overcome vulnerabilities.

Not less important we also are going to describe how we have implemented our application time-line and the functionalities above it such as time manipulation and annotations management. Among other different functionalities we also are going to describe how we perform stream composition, synchronization of our collaborative text editor and implementation of our instant messenger.

Lastly, we are going to describe the deployment of our solution in respect to the hardware that we have used and the software installed onto it.

\input{13_implementation_model.tex}
\input{14_implementation_signaling.tex}
\input{15_implementation_recording.tex}
\input{16_implementation_content.tex}
\input{17_implementation_deploy.tex}
\input{18_implementation_summary.tex}
